Parallelization of the Probabilistic Roadmap Method with GPU Acceleration

### 2016-07-27 paper/paper.tex
subtitle weg
qs qe, float2,.. schrift

\title{ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
\subtitle{High Performance Computing Project}
%\date{}
\author{Manuel Demmeler \and Manfred Liebmann}

@@

Now the movement of the robot can be described by a curve $q(t) \in \R^d, t\in[0,1]$, in the configuration space. 
The movement is collision-free, if at no time, any two polytopes of the robot intersect with each other or with the environment after being transformed into the world frame.
To apply the probabilistic roadmap algorithm, we express this setting by the indicator function
\begin{equation}
	I(q)=
	\begin{cases}
		0,	& \text{ for no collision of any polytope pair} \\
		1,	& \text{ if at least one collision occurs },
	\end{cases}
\end{equation}

@@

For the implementation of the kinematics and later the collision algorithm, a geometry library was written, based on the CUDA datatypes float2 for 2D vectors and float4 for 3D vectors.
To guarantee aligned access on the GPU, float4 is used instead of float3. 
Furthermore, as the precision of the geometric calculations is numerically not very critical, the usage of single precision instead of double precision did not cause problems.
The library contains host and device inline functions for common vector operations as shown in code \ref{code1}.
Operators like $+$ or $*$ are not overloaded to prevent unnecessary temporary variables.
Transformation matrices are implemented as float4 arrays with inline functions for common matrix operations. Again operators with temporaries were avoided, wherever possible.

@@

Code \ref{code4} exemplarily shows the class definition of the robot configuration space.
The geometry and robot data does not change during runtime
and is load\-ed once at the beginning and copied to the GPU by the 
\footnotesize
init
\normalsize 
function.
The indicator function gets as input a list of start and end points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if the indicator function is always 0 on the line in between.
\begin{equation}
	\mathrm{res[j]}=\begin{cases}
		%0	&\text{if $I(q)=0$ f.a. $q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q}$} \\
		0	&\text{if\ \ } \forall q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q} : I(q)=0 \\
		1	&\text{else}
	\end{cases}
\end{equation}
This computation is done by calling the kernel in Code \ref{code5}.
The start and end points are stored in structures of arrays defined by
\begin{equation}
\begin{aligned}
	q_{start}^j &= (q_s^j , q_s^{j+N_{buf}} , ... , q_s^{j+(d-1)*N_{buf}})^T\\
	q_{end}^j &= (q_e^j , q_e^{j+N_{buf}} , ... , q_e^{j+(d-1)*N_{buf}})^T
\end{aligned}
\end{equation}
As the PRM solver only needs to know, which start and end points can be connected by a line, the kernel delivers exactly this minimal necessary information. 
CPU and communication time is also saved by computing interpolating points directly on the GPU and not to exchange them.
Algorithms \ref{host} and \ref{kernel} show in pseudo code, how the work is split between host and device.

\begin{algorithm}\label{host}
	\KwData{qs, qe, N, offset}
	\KwResult{res}
	compute distances $d_j = \abs{q_{end}^j - q_{start}^j}$ \;
	$\mathrm{testnum[j]} = d_j/{\Delta q}$: number of threads for each line \;
	$\mathrm{testpos[j]} = \mathrm{thread\ displacements}$ \;
	reset res \;
	call kernel with $\sum_{j=0}^N \mathrm{testnum[j]}$ threads\;
	\caption{Host indicator function}
\end{algorithm}

\begin{algorithm}\label{kernel}
	\KwData{arrays qs, qe, testpos, testnum and threadindex}
	\KwResult{res}
	compute associated edge number $j$, defined by $testpos[j] \leq threadindex \le testpos[j]+testnum[j]$ \;
	$c = (\mathrm{threadindex - testpos[j]}) / (\mathrm{testnum[k]}-1) $ \;
	$q=c q_{start}^j + (1-c) q_{end}^j $ \;
	calculate robot kinematics $T_0(q), ..., T_d(q) $\;
	$\mathrm{res\_tmp} = 0; $ \;
	\For{all registered pairs pairs of polytopes $P, Q$}{
		$\mathrm{res\_tmp}=\mathrm{separating\_vector\_algorithm}(P, T_{i_P}, Q, T_{i_Q})$ \;
		\If{$\mathrm{res\_tmp} \neq 0$}{
			$\mathrm{res}[j]=\mathrm{res\_tmp}$ \;
			break \;
		}
	}
	\caption{Robot kernel: code for one thread}
\end{algorithm}

>>>>

\title{ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
%\subtitle{High Performance Computing Project}
%\date{}
\author{Manuel Demmeler \and Manfred Liebmann}

@@

Now the movement of the robot can be described by a curve $q(t) \in \R^d, t\in[0,1]$, in the configuration space. 
The movement is collision-free, if at no time, any two polytopes of the robot intersect with each other or with the environment after being transformed into the world frame.
To apply the probabilistic roadmap algorithm, we express this setting by the indicator function
\begin{equation}
	I(q)=
	\begin{cases}
		0,	& \text{ for no collision of any polytope pair} \\
		1,	& \text{ if at least one collision occurs }
	\end{cases}
	\; .
\end{equation}

@@

For the implementation of the kinematics and later the collision algorithm, a geometry library was written, based on the CUDA datatypes \texttt{float2} for 2D vectors and \texttt{float4} for 3D vectors.
To guarantee aligned access on the GPU, float4 is used instead of float3. 
Furthermore, as the precision of the geometric calculations is numerically not very critical, the usage of single precision instead of double precision did not cause problems.
The library contains host and device inline functions for common vector operations as shown in code \ref{code1}.
Operators like $+$ or $*$ are not overloaded to prevent unnecessary temporary variables.
Transformation matrices are implemented as float4 arrays with inline functions for common matrix operations. Again operators with temporaries were avoided, wherever possible.

@@

Code \ref{code4} exemplarily shows the class definition of the robot configuration space.
The geometry and robot data does not change during runtime
and is load\-ed once at the beginning and copied to the GPU by the 
\texttt{init}
function.
The indicator function gets as input a list of start and end points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if the indicator function is always 0 on the line in between.
\begin{equation}
	\mathrm{res[j]}=\begin{cases}
		%0	&\text{if $I(q)=0$ f.a. $q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q}$} \\
		0	&\text{if\ \ } \forall q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q} : I(q)=0 \\
		1	&\text{else}
	\end{cases}
\end{equation}
This computation is done by calling the kernel in Code \ref{code5}.
The start and end points are stored in structures of arrays defined by
\begin{equation}
\begin{aligned}
	q_{start}^j &= (q_s^j , q_s^{j+N_{buf}} , ... , q_s^{j+(d-1)*N_{buf}})^T\\
	q_{end}^j &= (q_e^j , q_e^{j+N_{buf}} , ... , q_e^{j+(d-1)*N_{buf}})^T
\end{aligned}
\end{equation}
As the PRM solver only needs to know, which start and end points can be connected by a line, the kernel delivers exactly this minimal necessary information. 
CPU and communication time is also saved by computing interpolating points directly on the GPU and not to exchange them.
Algorithms \ref{host} and \ref{kernel} show in pseudo code, how the work is split between host and device.

\begin{algorithm}\label{host}
	\KwData{$q_s$, $q_e$, N, $N_{buf}$}
	\KwResult{res}
	compute distances $d_j = \abs{q_{end}^j - q_{start}^j}$ \;
	$\mathrm{testnum[j]} = d_j/{\Delta q}$: number of threads for each line \;
	$\mathrm{testpos[j]} = \mathrm{thread\ displacements}$ \;
	reset res \;
	call kernel with $\sum_{j=0}^N \mathrm{testnum[j]}$ threads\;
	\caption{Host indicator function}
\end{algorithm}

\begin{algorithm}\label{kernel}
	\KwData{arrays $q_s$, $q_e$, testpos, testnum and threadindex}
	\KwResult{res}
	compute associated edge number $j$, defined by $testpos[j] \leq threadindex \le testpos[j]+testnum[j]$ \;
	$c = (\mathrm{threadindex - testpos[j]}) / (\mathrm{testnum[k]}-1) $ \;
	$q=c q_{start}^j + (1-c) q_{end}^j $ \;
	calculate robot kinematics $T_0(q), ..., T_d(q) $\;
	$\mathrm{res\_tmp} = 0; $ \;
	\For{all registered pairs pairs of polytopes $P, Q$}{
		$\mathrm{res\_tmp}=\mathrm{separating\_vector\_algorithm}(P, T_{i_P}, Q, T_{i_Q})$ \;
		\If{$\mathrm{res\_tmp} \neq 0$}{
			$\mathrm{res}[j]=\mathrm{res\_tmp}$ \;
			break \;
		}
	}
	\caption{Robot kernel: code for one thread}
\end{algorithm}
